---
title: "step011-IRIS_batch123_tertiary_analysis_exhaustion_markers"
author: "Ming"
date: "`r format(Sys.time(), '%Y_%m_%d_%H_%M')`"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(eval = FALSE)      ## no run
# knitr::opts_chunk$set(include = FALSE)   ## yes run, no code, no output
knitr::opts_chunk$set(echo = TRUE)         ## yes run, no code, yes output ( for figures )
knitr::opts_chunk$set(results = 'hide')    ## yes run, yes code, no output
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
'%ni%' = Negate('%in%')
# options(scipen = 999)       # decimals
# options(scipen = 0)         # sci notation
# options(max.print=999999)

dir.ref                 = "/Users/minghan/bioinfoproj/references"
dir.clin                = "/Users/minghan/GDrive_minghanpughlab/PughLabPMH/_projects/IRIS/_clin"
dir.data                = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/data"
dir.out.data            = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_data"
dir.out.figure.analysis = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_analysis"
dir.out.figure.qc       = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_qc"

PROJ = "IRIS"
PROJ_BATCH = "IRIS_batch123"

colours_n40 = 
  c("dodgerblue2", "#E31A1C", "green4", "#6A3D9A", "#FF7F00", 
    "gold1", "gray70", "#C1645C", "skyblue2", "#FB9A99", "palegreen2", 
    "#CAB2D6", "#FDBF6F", "#D46A43", "khaki2", "maroon", 
    "orchid1", "deeppink1", "blue1", "steelblue4", "darkturquoise", 
    "green1", "yellow4", "yellow3", "darkorange4", "brown",
    "#5C9A5C", "#CC8BAD", "#B8782A", "#546C9D", "#459E6F",
    
    "chocolate4", "chartreuse4", "burlywood", "deeppink4", "lightslateblue",
    "firebrick", "mintcream", "mediumpurple3", "springgreen4", "slateblue4", "pink")

library(Seurat)
```

# readRDS()

## all cells

```{r}
crescent.seurat.obj.singleR.encode.scmulti = readRDS(file.path(dir.out.data, paste0(PROJ_BATCH, ".crescent.seurat.obj.singleR.encode.all_metadata_attached.Rds")))
crescent.seurat.obj.singleR.encode.scmulti@meta.data %>% dplyr::select(dataset, Community_specificity_lean2) %>% table()
#                 Community_specificity_lean2
# dataset          no_GLIPHII_results tumor_associated_HomoSapiens tumor_associated_HomoSapiens_MDavis tumor_associated_MDavis viral_or_nonspecific
#   IRS_0004_A_003               2980                            0                                   5                      44                  123
#   IRS_0014_P_014               1507                            0                                   0                       6                    6
#   IRS_002_A_001                3514                            0                                   4                      84                  238
#   IRS_0023_P_023              10020                            1                                  35                     160                  130
#   IRS_0025_A_025                371                            0                                   1                       7                    3
#   IRS_0032_A_032                878                            0                                  15                       7                    7
#   IRS_0038_P_038               7772                            0                                  28                      74                   51
#   IRS_0055_P_055               7146                            1                                  24                     185                  105
#   IRS_0056_P_056                925                            0                                   4                      18                    9
#   IRS_008_P_008                 212                            0                                   1                       8                   14
#   IRS_015_A_015                8793                            0                                 200                     127                  121
```

## T cells re-integrated

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT = readRDS(file.path(dir.out.data, paste0(PROJ_BATCH, ".crescent.seurat.obj.T_cells.integrated.no_gdT.annotated.Rds")))
crescent.seurat.obj.T_cells.integrated.no_gdT # 34884 features across 12980 samples within 3 assays 

## attach metadata
crescent.seurat.obj.T_cells.integrated.no_gdT.meta.data = crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data
crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data = 
  crescent.seurat.obj.singleR.encode.scmulti@meta.data %>% 
  dplyr::filter(rownames(.) %in% (crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% rownames())) %>% 
  mutate(dataset = factor(dataset, levels = c("IRS_0004_A_003", "IRS_002_A_001", "IRS_0025_A_025", "IRS_0032_A_032", "IRS_015_A_015", 
                                              "IRS_0014_P_014", "IRS_0023_P_023", "IRS_0038_P_038", "IRS_0055_P_055", "IRS_0056_P_056", "IRS_008_P_008")))
crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data
```

# --------------------------------------------

# exhaustion markers

## sample_order

```{r}
sample_order = 
  c("IRS_0004_A_003", "IRS_002_A_001", "IRS_0025_A_025", "IRS_0032_A_032", "IRS_015_A_015", 
    "IRS_0014_P_014", "IRS_0023_P_023", "IRS_0038_P_038", "IRS_0055_P_055", "IRS_0056_P_056", "IRS_008_P_008")

sample_order.Tcell_num_gt500 = 
  c("IRS_0004_A_003", "IRS_002_A_001", "IRS_015_A_015", 
    "IRS_0023_P_023", "IRS_0038_P_038", "IRS_0055_P_055", "IRS_0056_P_056")
```

## markers

```{r}
immunomod.targets.genes.Tcell = 
  c("PDCD1", "TIGIT", "CD96", "CD226", "LAG3", "CCR8", "HAVCR2", "CTLA4", "PVRIG") 
immunomod.targets.genes.tumour = 
  c("CD274", "PVR", "NECTIN2")

CD226.specific.genes = 
  c("CD226", "TIGIT", "PVRIG", "CD96", "NECTIN2", "PVR", "NECTIN3", "NECTIN1", "NECTIN4")

markers_all = c("PDCD1", "TIGIT", "LAG3", "CD96", "CD274", "PVR", "PVRIG", "CD226", 
                "HAVCR2", "CCR8", "NECTIN1", "NECTIN2", "NECTIN3", "NECTIN4")
```

## plots - all T cells

### DotPlot()

```{r}
Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT, 
                features = immunomod.targets.genes.Tcell, 
                assay = "SCT", 
                group.by = "dataset",
                cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_all.DotPlot.allSamples_v2.png")), 
#        device = "png", width = 6, height = 4, units = "in")

Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT, 
                features = immunomod.targets.genes.Tcell, 
                assay = "SCT", 
                group.by = "dataset", 
                cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order.Tcell_num_gt500)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_all.DotPlot.samplesTcells_gt500_v2.png")), 
#        device = "png", width = 6, height = 4, units = "in")
```

### boxplot

#### per cell

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.metadata = crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data
crescent.seurat.obj.T_cells.integrated.no_gdT.metadata

# crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data = 
#   crescent.seurat.obj.T_cells.integrated.no_gdT.metadata %>% 
#   rownames_to_column(var="sample_barcode") %>% 
#   left_join(., clin.df %>% mutate(Patient_ID = Patient_ID %>% gsub("-","_", .)), 
#             by = c("dataset"="Patient_ID")) # %>% dplyr::filter(is.na(Deceased))

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df = 
  crescent.seurat.obj.T_cells.integrated.no_gdT@assays$SCT@data[immunomod.targets.genes.Tcell, ] %>% data.frame()

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 = 
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df %>% 
  rownames_to_column(var="SYMBOL") %>% 
  pivot_longer(cols = -SYMBOL, names_to = "sample_barcode", values_to = "SCTdata") %>% 
  # mutate(sample = sample_barcode %>% gsub(".{17}$", "", .)) %>% 
  left_join(., crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% rownames_to_column(var="sample_barcode"), by=c("sample_barcode"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_all.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% 
  print.data.frame()
#   SYMBOL   Acquired   Primary
# 1   CCR8 0.06983218 0.1489022
# 2  CD226 0.20084178 0.2071352
# 3   CD96 0.97152595 1.0063079
# 4  CTLA4 0.57313144 0.7610295
# 5 HAVCR2 0.36941195 0.3013368
# 6   LAG3 1.02780347 1.0469986
# 7  PDCD1 0.47209279 0.4136221
# 8  PVRIG 0.34282741 0.3152963
# 9  TIGIT 0.73759765 0.8307842
```

#### pseudo-bulk

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df3 =
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::group_by(dataset, SYMBOL) %>% 
  summarise(SCTdata = mean(SCTdata)) %>% 
  left_join(., 
            crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% 
              dplyr::select(dataset, starts_with("clin_")) %>% 
              distinct() %>% remove_rownames(), 
            by=c("dataset"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df3

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df3 %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_all.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df3 %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% 
  print.data.frame()
#   SYMBOL  Acquired   Primary
# 1   CCR8 0.1115888 0.1416402
# 2  CD226 0.1872667 0.1933002
# 3   CD96 0.9583865 0.9964493
# 4  CTLA4 0.6978032 0.7239057
# 5 HAVCR2 0.2750779 0.2409164
# 6   LAG3 0.8353138 0.8006743
# 7  PDCD1 0.4272648 0.3645690
# 8  PVRIG 0.3296104 0.3390799
# 9  TIGIT 0.7897905 0.6420451
```

## plots - tumor_assoc T cells

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data$Community_specificity_lean2 %>% table()

## tumor associated
crescent.seurat.obj.T_cells.integrated.no_gdT.t_assoc = subset(crescent.seurat.obj.T_cells.integrated.no_gdT, subset = Community_specificity_lean == "tumor_associated")

## Human
crescent.seurat.obj.T_cells.integrated.no_gdT.human = subset(crescent.seurat.obj.T_cells.integrated.no_gdT, subset = Community_specificity_lean2 == "tumor_associated_HomoSapiens")
crescent.seurat.obj.T_cells.integrated.no_gdT.human # 34884 features across 2 samples within 3 assays 

## MDavis
crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis = subset(crescent.seurat.obj.T_cells.integrated.no_gdT, subset = Community_specificity_lean2 == "tumor_associated_MDavis")
crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis # 34884 features across 618 samples within 3 assays 

## Human_MDavis
crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis = subset(crescent.seurat.obj.T_cells.integrated.no_gdT, subset = Community_specificity_lean2 == "tumor_associated_HomoSapiens_MDavis")
crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis # 34884 features across 249 samples within 3 assays 
```

### DotPlot()

#### tumor associated

```{r}
Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.t_assoc, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.DotPlot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 4, units = "in")

Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.t_assoc, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order.Tcell_num_gt500)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

#### MDavis

```{r}
Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.DotPlot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 4, units = "in")

Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order.Tcell_num_gt500)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

#### Human_MDavis

```{r}
Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.DotPlot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 4, units = "in")

Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order.Tcell_num_gt500)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

### boxplot

#### per cell

##### tumor associated

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc = 
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.t_assoc@meta.data %>% rownames()))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc$sample_barcode %>% unique() %>% length() # [1] 869

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean")
```

##### MDavis

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis = 
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis@meta.data %>% rownames()))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis$sample_barcode %>% unique() %>% length() # [1] 618

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_mdavis.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean")
```

##### Human_MDavis

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis = 
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis@meta.data %>% rownames()))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis$sample_barcode %>% unique() %>% length() # [1] 249

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_human_mdavis.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean")
```

#### pseudo-bulk

##### tumor associated

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc.pseudobulk =
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.t_assoc@meta.data %>% rownames())) %>% 
  dplyr::group_by(dataset, SYMBOL) %>% 
  summarise(SCTdata = mean(SCTdata)) %>% 
  left_join(., 
            crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% 
              dplyr::select(dataset, starts_with("clin_")) %>% 
              distinct() %>% remove_rownames(), 
            by=c("dataset"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc.pseudobulk

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc.pseudobulk %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.tumour_assoc.pseudobulk %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% print.data.frame()
#   SYMBOL   Acquired   Primary
# 1   CCR8 0.09728914 0.0975103
# 2  CD226 0.18549383 0.1877208
# 3   CD96 0.98972327 1.0643769
# 4  CTLA4 0.81547981 0.6425010
# 5 HAVCR2 0.40461654 0.2096404
# 6   LAG3 0.91831852 0.8240537
# 7  PDCD1 0.42461112 0.2973220
# 8  PVRIG 0.37483780 0.2973290
# 9  TIGIT 0.92313415 0.6458892
```

##### MDavis

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis.pseudobulk =
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.mdavis@meta.data %>% rownames())) %>% 
  dplyr::group_by(dataset, SYMBOL) %>% 
  summarise(SCTdata = mean(SCTdata)) %>% 
  left_join(., 
            crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% 
              dplyr::select(dataset, starts_with("clin_")) %>% 
              distinct() %>% remove_rownames(), 
            by=c("dataset"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis.pseudobulk

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis.pseudobulk %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_mdavis.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.mdavis.pseudobulk %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% print.data.frame()
#   SYMBOL  Acquired   Primary
# 1   CCR8 0.1210464 0.1065839
# 2  CD226 0.1882426 0.1504379
# 3   CD96 0.9322761 1.0146406
# 4  CTLA4 0.7317719 0.6543424
# 5 HAVCR2 0.2930740 0.2019039
# 6   LAG3 0.6695068 0.8123101
# 7  PDCD1 0.2854271 0.2984325
# 8  PVRIG 0.3460723 0.2946358
# 9  TIGIT 0.8444566 0.6311071
```

##### Human_MDavis

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis.pseudobulk =
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.human_mdavis@meta.data %>% rownames())) %>% 
  dplyr::group_by(dataset, SYMBOL) %>% 
  summarise(SCTdata = mean(SCTdata)) %>% 
  left_join(., 
            crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% 
              dplyr::select(dataset, starts_with("clin_")) %>% 
              distinct() %>% remove_rownames(), 
            by=c("dataset"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis.pseudobulk

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis.pseudobulk %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_human_mdavis.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.human_mdavis.pseudobulk %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% print.data.frame()
#   SYMBOL   Acquired   Primary
# 1   CCR8 0.01083879 0.0602127
# 2  CD226 0.17548425 0.5346927
# 3   CD96 1.05410257 1.3704452
# 4  CTLA4 0.74425638 0.4924700
# 5 HAVCR2 0.27384072 0.1867475
# 6   LAG3 1.12998171 0.7533328
# 7  PDCD1 0.51882582 0.2736502
# 8  PVRIG 0.45784112 0.3064454
# 9  TIGIT 1.07431603 0.7850246
```




## plots - viral_or_nonspecific T cells

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc = subset(crescent.seurat.obj.T_cells.integrated.no_gdT, subset = Community_specificity_lean == "viral_or_nonspecific")
crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc@meta.data
```

### DotPlot()

```{r}
Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_non_tumour_assoc.DotPlot.allSamples.png")), 
#        device = "png", width = 7, height = 4, units = "in")

Seurat::DotPlot(object = crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc, features = immunomod.targets.genes.Tcell, assay = "SCT", group.by = "dataset", cols = colours_n40) +
  scale_y_discrete(limits = rev(sample_order.Tcell_num_gt500)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

### boxplot

#### per cell

##### tumor associated

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc = 
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc@meta.data %>% rownames()))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc$sample_barcode %>% unique() %>% length() # [1] 721

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% print.data.frame()
#   SYMBOL   Acquired   Primary
# 1   CCR8 0.02786379 0.1985414
# 2  CD226 0.14047657 0.1464414
# 3   CD96 1.07560683 0.9447712
# 4  CTLA4 0.45510651 0.8643146
# 5 HAVCR2 0.35788216 0.3299338
# 6   LAG3 1.21974135 1.0922771
# 7  PDCD1 0.50776829 0.4282011
# 8  PVRIG 0.47413560 0.3786951
# 9  TIGIT 0.80481541 0.9082835
```

#### pseudo-bulk

##### tumor associated

```{r}
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc.pseudobulk =
  crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.df2 %>% 
  dplyr::filter(sample_barcode %in% c(crescent.seurat.obj.T_cells.integrated.no_gdT.non_tumour_assoc@meta.data %>% rownames())) %>% 
  dplyr::group_by(dataset, SYMBOL) %>% 
  summarise(SCTdata = mean(SCTdata)) %>% 
  left_join(., 
            crescent.seurat.obj.T_cells.integrated.no_gdT@meta.data %>% 
              dplyr::select(dataset, starts_with("clin_")) %>% 
              distinct() %>% remove_rownames(), 
            by=c("dataset"))
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc.pseudobulk

crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc.pseudobulk %>% 
  ggplot(., aes(x = clin_ResistanceType, y = SCTdata)) + 
	geom_boxplot(notch = F, outlier.shape = NA) + 
  geom_jitter(shape=16, size=0.3, position=position_jitter(0.2)) +
  ggpubr::stat_compare_means(method="wilcox", size = 3, comparisons=list(c(1,2))) +
	scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) + 
  facet_wrap( ~ SYMBOL, nrow = 3, scales = "free") +
  theme_classic()
# ggsave(file.path(dir.out.figure.analysis, paste0(PROJ_BATCH, ".seurat.obj.exhaustMarkers.Tcells_tumour_assoc.boxplot.allSamples_newREF.png")), 
#        device = "png", width = 7, height = 7, units = "in", dpi = 300, bg = "white")

## get means
crescent.seurat.obj.T_cells.integrated.no_gdT.SCTdata.im_target_genes_Tcell.non_tumour_assoc.pseudobulk %>% 
  group_by(SYMBOL, clin_ResistanceType) %>% 
  summarize(SCTdata_mean = mean(SCTdata)) %>% 
  pivot_wider(id_cols = SYMBOL, names_from = "clin_ResistanceType", values_from = "SCTdata_mean") %>% print.data.frame()
#   SYMBOL  Acquired   Primary
# 1   CCR8 0.2139558 0.1478142
# 2  CD226 0.1581311 0.1559208
# 3   CD96 1.0327221 0.9801782
# 4  CTLA4 0.8074128 0.7964050
# 5 HAVCR2 0.2743245 0.2223705
# 6   LAG3 0.9741598 0.8088163
# 7  PDCD1 0.4769683 0.3599690
# 8  PVRIG 0.4059572 0.3826182
# 9  TIGIT 0.8751769 0.7373020
```

<br>

# EOF
