---
title: "2023_08_01_1159-IRIS_batch123_scTCR_BCR_wrangle_metadata"
author: "Ming"
date: "`r format(Sys.time(), '%Y_%m_%d_%H_%M')`"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(eval = FALSE)      ## no run
# knitr::opts_chunk$set(include = FALSE)   ## yes run, no code, no output
knitr::opts_chunk$set(echo = TRUE)         ## yes run, no code, yes output ( for figures )
knitr::opts_chunk$set(results = 'hide')    ## yes run, yes code, no output
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
'%ni%' = Negate('%in%')
options(scipen = 999)       # decimals
# options(scipen = 0)         # sci notation
# options(max.print=999999)

dir.data                = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/data"
dir.out.data            = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_data"
dir.out.figure.analysis = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_analysis"
dir.out.figure.qc       = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_qc"

dir.data.scTCR          = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_for_immunarch_scTCR"
dir.data.scBCR          = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_for_immunarch_scBCR"

PROJ_BATCH = "IRIS_batch123"

colours_n30 = 
  c("dodgerblue2", "#E31A1C", "green4", "#6A3D9A", "#FF7F00", 
    "gold1", "gray70", "#C1645C", "skyblue2", "#FB9A99", "palegreen2", 
    "#CAB2D6", "#FDBF6F", "#D46A43", "khaki2", "maroon", 
    "orchid1", "deeppink1", "blue1", "steelblue4", "darkturquoise", 
    "green1", "yellow4", "yellow3", "darkorange4", "brown",
    "#5C9A5C", "#CC8BAD", "#B8782A", "#546C9D", "#459E6F")

expanded_cutoff = 1
cloneTypes = c(Rare = 1e-04, Small = 0.001, Medium = 0.01, Large = 0.1, Hyperexpanded = 1)
# cloneTypes
```

# read in data

## scTCR

```{r}
scTCR.im.prop_recalcd.TRA.list = readRDS(file.path(dir.out.data, "scTCR.im.prop_recalcd.TRA.list.Rds"))
scTCR.im.prop_recalcd.TRB.list = readRDS(file.path(dir.out.data, "scTCR.im.prop_recalcd.TRB.list.Rds"))
```

## scBCR

```{r}
scBCR.im.prop_recalcd.IGH.list = readRDS(file.path(dir.out.data, "scBCR.im.prop_recalcd.IGH.list.Rds"))
scBCR.im.prop_recalcd.IGK.list = readRDS(file.path(dir.out.data, "scBCR.im.prop_recalcd.IGK.list.Rds"))
scBCR.im.prop_recalcd.IGL.list = readRDS(file.path(dir.out.data, "scBCR.im.prop_recalcd.IGL.list.Rds"))
```

# --------------------------------------------

# get scTCR barcodes

## wrangle TRA

```{r}
scTCR.im.prop_recalcd.TRA.cell_barcode.df =
  scTCR.im.prop_recalcd.TRA.list %>% 
  bind_rows(., .id = "sample") %>% 
  dplyr::select(sample, Barcode, CDR3.aa, V.name, J.name, Clones, Proportion) %>%
  tidyr::separate_longer_delim(c(Barcode), delim = ";") %>% 
  mutate(sample2 = sample %>% gsub("_filtered_contig_annotations_TRA", "", .),
         Proportion = Proportion %>% round(5)) %>% 
  dplyr::rename("sample_TRA"="sample",
                "CDR3.aa_TRA"="CDR3.aa",
                "V.name_TRA"="V.name",
                "J.name_TRA"="J.name",
                "Clones_TRA"="Clones",
                "Proportion_TRA"="Proportion") %>% 
  unite(col = "sample2_Barcode", sample2, Barcode, sep = "__", remove = T) %>% 
  mutate(TRA_expanded = if_else(Clones_TRA > expanded_cutoff, "expanded", "not_expanded"),
         TRA_cloneSize = case_when(Proportion_TRA >= 0.9    & Proportion_TRA < 1     ~ "Hyperexpanded",
                                   Proportion_TRA >= 0.1    & Proportion_TRA < 0.9   ~ "Large",
                                   Proportion_TRA >= 0.01   & Proportion_TRA < 0.1   ~ "Medium",
                                   Proportion_TRA >= 0.001  & Proportion_TRA < 0.01  ~ "Small",
                                   Proportion_TRA >= 0.0001 & Proportion_TRA < 0.001 ~ "Rare",
                                   TRUE ~ "ERROR")) %>% 
  arrange(sample2_Barcode)
scTCR.im.prop_recalcd.TRA.cell_barcode.df %>% dim() # [1] 16355     9
# scTCR.im.prop_recalcd.TRA.cell_barcode.df
```

## wrangle TRB

```{r}
scTCR.im.prop_recalcd.TRB.cell_barcode.df =
  scTCR.im.prop_recalcd.TRB.list %>% 
  bind_rows(., .id = "sample") %>% 
  dplyr::select(sample, Barcode, CDR3.aa, V.name, J.name, Clones, Proportion) %>%
  tidyr::separate_longer_delim(c(Barcode), delim = ";") %>% 
  mutate(sample2 = sample %>% gsub("_filtered_contig_annotations_TRB", "", .),
         Proportion = Proportion %>% round(5)) %>% 
  dplyr::rename("sample_TRB"="sample",
                "CDR3.aa_TRB"="CDR3.aa",
                "V.name_TRB"="V.name",
                "J.name_TRB"="J.name",
                "Clones_TRB"="Clones",
                "Proportion_TRB"="Proportion") %>% 
  unite(col = "sample2_Barcode", sample2, Barcode, sep = "__", remove = T) %>% 
  mutate(TRB_expanded = if_else(Clones_TRB > expanded_cutoff, "expanded", "not_expanded"),
         TRB_cloneSize = case_when(Proportion_TRB >= 0.9    & Proportion_TRB < 1     ~ "Hyperexpanded",
                                   Proportion_TRB >= 0.1    & Proportion_TRB < 0.9   ~ "Large",
                                   Proportion_TRB >= 0.01   & Proportion_TRB < 0.1   ~ "Medium",
                                   Proportion_TRB >= 0.001  & Proportion_TRB < 0.01  ~ "Small",
                                   Proportion_TRB >= 0.0001 & Proportion_TRB < 0.001 ~ "Rare",
                                   TRUE ~ "ERROR")) %>% 
  arrange(sample2_Barcode)
scTCR.im.prop_recalcd.TRB.cell_barcode.df %>% dim() # [1] 16986     9
# scTCR.im.prop_recalcd.TRB.cell_barcode.df %>% names() %>% cat(., sep = ", ")
```

## full_join() TRA and TRB

```{r}
scTCR.im.prop_recalcd.TRAandB.cell_barcode.df = 
  full_join(scTCR.im.prop_recalcd.TRA.cell_barcode.df, 
            scTCR.im.prop_recalcd.TRB.cell_barcode.df,
            by=c("sample2_Barcode"="sample2_Barcode")) %>% 
  dplyr::select(sample2_Barcode, 
                CDR3.aa_TRA, V.name_TRA, J.name_TRA, Clones_TRA, Proportion_TRA, TRA_expanded, TRA_cloneSize, 
                CDR3.aa_TRB, V.name_TRB, J.name_TRB, Clones_TRB, Proportion_TRB, TRB_expanded, TRB_cloneSize)
scTCR.im.prop_recalcd.TRAandB.cell_barcode.df %>% dim() # [1] 19134     15
# scTCR.im.prop_recalcd.TRAandB.cell_barcode.df
```

## collapse {TRA;TRB}

```{r}
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed = 
  scTCR.im.prop_recalcd.TRAandB.cell_barcode.df %>% 
  group_by(sample2_Barcode) %>%
  summarize(CDR3.aa_TRA    = paste(CDR3.aa_TRA,    collapse = ";"), 
            V.name_TRA     = paste(V.name_TRA,     collapse = ";"), 
            J.name_TRA     = paste(J.name_TRA,     collapse = ";"), 
            Clones_TRA     = paste(Clones_TRA,     collapse = ";"), 
            Proportion_TRA = paste(Proportion_TRA, collapse = ";"), 
            TRA_expanded   = paste(TRA_expanded,   collapse = ";"), 
            TRA_cloneSize  = paste(TRA_cloneSize,  collapse = ";"), 
            
            CDR3.aa_TRB    = paste(CDR3.aa_TRB,    collapse = ";"), 
            V.name_TRB     = paste(V.name_TRB,     collapse = ";"), 
            J.name_TRB     = paste(J.name_TRB,     collapse = ";"), 
            Clones_TRB     = paste(Clones_TRB,     collapse = ";"), 
            Proportion_TRB = paste(Proportion_TRB, collapse = ";"), 
            TRB_expanded   = paste(TRB_expanded,   collapse = ";"), 
            TRB_cloneSize  = paste(TRB_cloneSize,  collapse = ";")) %>% 
  mutate(across(everything(), .fns = ~ gsub("^NA;.*NA$", "NA", .x)))
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dim() # [1] 16936    15
# scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed
```

## sanity checks

```{r}
## cells expressing max 1 TCR
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dplyr::filter(!str_detect(CDR3.aa_TRA, ";")) %>% nrow() # [1] 15032
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dplyr::filter(!str_detect(CDR3.aa_TRB, ";")) %>% nrow() # [1] 15032

## cells express both TRA and TRB
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dplyr::filter(str_detect(CDR3.aa_TRB, "^NA$")) # [1] 379 <-- no TRB
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dplyr::filter(str_detect(CDR3.aa_TRA, "^NA$")) # [1] 2203 <-- no TRA
scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% dplyr::filter(!str_detect(CDR3.aa_TRA, "^NA$") & !str_detect(CDR3.aa_TRB, "^NA$")) # [1] 14354 <-- both TRA and TRB
```

## get scTCR.metadata

```{r}
scTCR.metadata = 
  scTCR.im.prop_recalcd.TRAandB.cell_barcode.collapsed %>% 
  mutate(Tcell = "yesTCR",
         TRA_num = case_when(!str_detect(CDR3.aa_TRA, ";") ~ "TRA_single",
                              str_detect(CDR3.aa_TRA, ";") ~ "TRA_multiple"), 
         TRB_num = case_when(!str_detect(CDR3.aa_TRB, ";") ~ "TRB_single",
                              str_detect(CDR3.aa_TRB, ";") ~ "TRB_multiple"),
         TR_A_B = case_when(str_detect(CDR3.aa_TRB, "^NA$") ~ "TRA_only",
                            str_detect(CDR3.aa_TRA, "^NA$") ~ "TRB_only",
                            TRUE ~ "TRA_and_TRB"))
scTCR.metadata %>% dim() # [1] 16936    19
```

## saveRDS()

```{r}
# scTCR.metadata %>% saveRDS(., file.path(dir.out.data, "scTCR.metadata.Rds"))
```

## inspect metadata

```{r}
scTCR.metadata$TRA_expanded  %>% table() %>% as.data.frame()
scTCR.metadata$TRA_cloneSize %>% table() %>% as.data.frame()
scTCR.metadata$TRA_num       %>% table() %>% as.data.frame()

scTCR.metadata$TRB_expanded  %>% table() %>% as.data.frame()
scTCR.metadata$TRB_cloneSize %>% table() %>% as.data.frame()
scTCR.metadata$TRB_num       %>% table() %>% as.data.frame()

scTCR.metadata$TR_A_B        %>% table() %>% as.data.frame()
```

# --------------------------------------------

# get scBCR barcodes

## wrangle IGH

```{r}
scTCR.im.prop_recalcd.IGH.cell_barcode.df =
  scBCR.im.prop_recalcd.IGH.list %>% 
  bind_rows(., .id = "sample") %>% 
  dplyr::select(sample, Barcode, CDR3.aa, V.name, J.name, Clones, Proportion) %>%
  tidyr::separate_longer_delim(c(Barcode), delim = ";") %>% 
  mutate(sample2 = sample %>% gsub("_filtered_contig_annotations_IGH", "", .),
         Proportion = Proportion %>% round(5)) %>% 
  dplyr::rename("sample_IGH"="sample",
                "CDR3.aa_IGH"="CDR3.aa",
                "V.name_IGH"="V.name",
                "J.name_IGH"="J.name",
                "Clones_IGH"="Clones",
                "Proportion_IGH"="Proportion") %>% 
  unite(col = "sample2_Barcode", sample2, Barcode, sep = "__", remove = T) %>% 
  mutate(IGH_expanded = if_else(Clones_IGH > expanded_cutoff, "expanded", "not_expanded"),
         IGH_cloneSize = case_when(Proportion_IGH >= 0.9    & Proportion_IGH < 1     ~ "Hyperexpanded",
                                   Proportion_IGH >= 0.1    & Proportion_IGH < 0.9   ~ "Large",
                                   Proportion_IGH >= 0.01   & Proportion_IGH < 0.1   ~ "Medium",
                                   Proportion_IGH >= 0.001  & Proportion_IGH < 0.01  ~ "Small",
                                   Proportion_IGH >= 0.0001 & Proportion_IGH < 0.001 ~ "Rare",
                                   TRUE ~ "ERROR")) %>% 
  arrange(sample2_Barcode)
scTCR.im.prop_recalcd.IGH.cell_barcode.df %>% dim() # [1] 1302     9
# scTCR.im.prop_recalcd.IGH.cell_barcode.df
```

## wrangle IGK

```{r}
scTCR.im.prop_recalcd.IGK.cell_barcode.df =
  scBCR.im.prop_recalcd.IGK.list %>% 
  bind_rows(., .id = "sample") %>% 
  dplyr::select(sample, Barcode, CDR3.aa, V.name, J.name, Clones, Proportion) %>%
  tidyr::separate_longer_delim(c(Barcode), delim = ";") %>% 
  mutate(sample2 = sample %>% gsub("_filtered_contig_annotations_IGK", "", .),
         Proportion = Proportion %>% round(5)) %>% 
  dplyr::rename("sample_IGK"="sample",
                "CDR3.aa_IGK"="CDR3.aa",
                "V.name_IGK"="V.name",
                "J.name_IGK"="J.name",
                "Clones_IGK"="Clones",
                "Proportion_IGK"="Proportion") %>% 
  unite(col = "sample2_Barcode", sample2, Barcode, sep = "__", remove = T) %>% 
  mutate(IGK_expanded = if_else(Clones_IGK > expanded_cutoff, "expanded", "not_expanded"),
         IGK_cloneSize = case_when(Proportion_IGK >= 0.9    & Proportion_IGK < 1     ~ "Hyperexpanded",
                                   Proportion_IGK >= 0.1    & Proportion_IGK < 0.9   ~ "Large",
                                   Proportion_IGK >= 0.01   & Proportion_IGK < 0.1   ~ "Medium",
                                   Proportion_IGK >= 0.001  & Proportion_IGK < 0.01  ~ "Small",
                                   Proportion_IGK >= 0.0001 & Proportion_IGK < 0.001 ~ "Rare",
                                   TRUE ~ "ERROR")) %>% 
  arrange(sample2_Barcode)
scTCR.im.prop_recalcd.IGK.cell_barcode.df %>% dim() # [1] 1010     9
# scTCR.im.prop_recalcd.IGK.cell_barcode.df
```

## wrangle IGL

```{r}
scTCR.im.prop_recalcd.IGL.cell_barcode.df =
  scBCR.im.prop_recalcd.IGL.list %>% 
  bind_rows(., .id = "sample") %>% 
  dplyr::select(sample, Barcode, CDR3.aa, V.name, J.name, Clones, Proportion) %>%
  tidyr::separate_longer_delim(c(Barcode), delim = ";") %>% 
  mutate(sample2 = sample %>% gsub("_filtered_contig_annotations_IGL", "", .),
         Proportion = Proportion %>% round(5)) %>% 
  dplyr::rename("sample_IGL"="sample",
                "CDR3.aa_IGL"="CDR3.aa",
                "V.name_IGL"="V.name",
                "J.name_IGL"="J.name",
                "Clones_IGL"="Clones",
                "Proportion_IGL"="Proportion") %>% 
  unite(col = "sample2_Barcode", sample2, Barcode, sep = "__", remove = T) %>% 
  mutate(IGL_expanded = if_else(Clones_IGL > expanded_cutoff, "expanded", "not_expanded"),
         IGL_cloneSize = case_when(Proportion_IGL >= 0.9    & Proportion_IGL < 1     ~ "Hyperexpanded",
                                   Proportion_IGL >= 0.1    & Proportion_IGL < 0.9   ~ "Large",
                                   Proportion_IGL >= 0.01   & Proportion_IGL < 0.1   ~ "Medium",
                                   Proportion_IGL >= 0.001  & Proportion_IGL < 0.01  ~ "Small",
                                   Proportion_IGL >= 0.0001 & Proportion_IGL < 0.001 ~ "Rare",
                                   TRUE ~ "ERROR")) %>% 
  arrange(sample2_Barcode)
scTCR.im.prop_recalcd.IGL.cell_barcode.df %>% dim() # [1] 538     9
# scTCR.im.prop_recalcd.IGL.cell_barcode.df %>% names() %>% cat(., sep = ", ")
```

## full_join IGH and IGK and IGL

```{r}
scTCR.im.prop_recalcd.IGH_K_L.cell_barcode.df = 
  plyr::join_all(list(scTCR.im.prop_recalcd.IGH.cell_barcode.df, 
                      scTCR.im.prop_recalcd.IGK.cell_barcode.df,
                      scTCR.im.prop_recalcd.IGL.cell_barcode.df), 
                 by=c("sample2_Barcode"), type = "full") %>% 
  dplyr::select(sample2_Barcode, 
                CDR3.aa_IGH, V.name_IGH, J.name_IGH, Clones_IGH, Proportion_IGH, IGH_expanded, IGH_cloneSize,                 
                CDR3.aa_IGK, V.name_IGK, J.name_IGK, Clones_IGK, Proportion_IGK, IGK_expanded, IGK_cloneSize, 
                CDR3.aa_IGL, V.name_IGL, J.name_IGL, Clones_IGL, Proportion_IGL, IGL_expanded, IGL_cloneSize)
scTCR.im.prop_recalcd.IGH_K_L.cell_barcode.df %>% dim() # [1] 1556     22
# scTCR.im.prop_recalcd.IGH_K_L.cell_barcode.df
```

## collapse {IGH;IGK;IGL}

```{r}
scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed =
  scTCR.im.prop_recalcd.IGH_K_L.cell_barcode.df %>% 
  group_by(sample2_Barcode) %>%
  summarize(CDR3.aa_IGH    = paste(CDR3.aa_IGH,    collapse = ";"), 
            V.name_IGH     = paste(V.name_IGH,     collapse = ";"), 
            J.name_IGH     = paste(J.name_IGH,     collapse = ";"), 
            Clones_IGH     = paste(Clones_IGH,     collapse = ";"), 
            Proportion_IGH = paste(Proportion_IGH, collapse = ";"), 
            IGH_expanded   = paste(IGH_expanded,   collapse = ";"), 
            IGH_cloneSize  = paste(IGH_cloneSize,  collapse = ";"), 
            
            CDR3.aa_IGK    = paste(CDR3.aa_IGK,    collapse = ";"), 
            V.name_IGK     = paste(V.name_IGK,     collapse = ";"), 
            J.name_IGK     = paste(J.name_IGK,     collapse = ";"), 
            Clones_IGK     = paste(Clones_IGK,     collapse = ";"), 
            Proportion_IGK = paste(Proportion_IGK, collapse = ";"), 
            IGK_expanded   = paste(IGK_expanded,   collapse = ";"), 
            IGK_cloneSize  = paste(IGK_cloneSize,  collapse = ";"), 
            
            CDR3.aa_IGL    = paste(CDR3.aa_IGL,    collapse = ";"), 
            V.name_IGL     = paste(V.name_IGL,     collapse = ";"), 
            J.name_IGL     = paste(J.name_IGL,     collapse = ";"), 
            Clones_IGL     = paste(Clones_IGL,     collapse = ";"), 
            Proportion_IGL = paste(Proportion_IGL, collapse = ";"), 
            IGL_expanded   = paste(IGL_expanded,   collapse = ";"), 
            IGL_cloneSize  = paste(IGL_cloneSize,  collapse = ";")) %>% 
  mutate(across(everything(), .fns = ~ gsub("^NA;.*NA$", "NA", .x)))
scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed %>% dim() # [1] 1542   22
# scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed
```

## sanity checks

```{r}
## cells expressing max 1 BCR
scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed %>% dplyr::filter(str_detect(CDR3.aa_IGH, ";", negate = T)) %>% nrow() # [1] 1532
scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed %>% dplyr::filter(str_detect(CDR3.aa_IGK, ";", negate = T)) %>% nrow() # [1] 1534
scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed %>% dplyr::filter(str_detect(CDR3.aa_IGL, ";", negate = T)) %>% nrow() # [1] 1539

## cells express IGH + IGK or IGL
temp1 = scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed
temp1 %>% dplyr::filter(!str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$")) # [1] 6   <-- only IGH
temp1 %>% dplyr::filter( str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$")) # [1] 149 <-- only IGK
temp1 %>% dplyr::filter( str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$")) # [1] 94  <-- only IGL
temp1 %>% dplyr::filter( str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$")) # [1] 0   <-- only IGK + IGL
temp1 %>% dplyr::filter(!str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$")) # [1] 3   <-- IGH + IGK + IGL
temp1 %>% dplyr::filter(!str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$")) # [1] 851 <-- with IGH + IGK
temp1 %>% dplyr::filter(!str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$")) # [1] 439 <-- with IGH + IGL

94 + 6 + 149 + 0 + 3 + 851 + 439 # [1] 1542 <-- matches total B cells
```

## get scBCR.metadata

```{r}
scBCR.metadata =
  scBCR.im.prop_recalcd.IGH_K_L.cell_barcode.collapsed %>% 
  mutate(Bcell = "yesBCR",
         IGH_num = case_when(!str_detect(CDR3.aa_IGH, ";") ~ "IGH_single",
                              str_detect(CDR3.aa_IGH, ";") ~ "IGH_multiple"),
         IGK_num = case_when(!str_detect(CDR3.aa_IGK, ";") ~ "IGK_single",
                              str_detect(CDR3.aa_IGK, ";") ~ "IGK_multiple"),
         IGL_num = case_when(!str_detect(CDR3.aa_IGL, ";") ~ "IGL_single",
                              str_detect(CDR3.aa_IGL, ";") ~ "IGL_multiple"),
         IGH_K_L = case_when(
           !str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$") ~ "IGH_only", 
            str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$") ~ "IGK_only", 
            str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$") ~ "IGL_only", 
            str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$") ~ "IGK_IGLonly", 
           !str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$") ~ "IGH_IGK_IGL", 
           !str_detect(CDR3.aa_IGH, "^NA$") & !str_detect(CDR3.aa_IGK, "^NA$") &  str_detect(CDR3.aa_IGL, "^NA$") ~ "IGH_IGK", 
           !str_detect(CDR3.aa_IGH, "^NA$") &  str_detect(CDR3.aa_IGK, "^NA$") & !str_detect(CDR3.aa_IGL, "^NA$") ~ "IGH_IGL", 
           TRUE ~ "ERROR"))
scBCR.metadata %>% dim() # [1] 1542   27
```

## saveRDS()

```{r}
# scBCR.metadata %>% saveRDS(., file.path(dir.out.data, "scBCR.metadata.Rds"))
```

## inspect metadata

```{r}
scBCR.metadata$IGH_num %>% table() # IGH_multiple-10   IGH_single-1532
scBCR.metadata$IGK_num %>% table() # IGK_multiple-8    IGK_single-1534
scBCR.metadata$IGL_num %>% table() # IGL_multiple-3    IGL_single-1539

scBCR.metadata$IGH_K_L %>% table()
#    IGH_IGK IGH_IGK_IGL     IGH_IGL    IGH_only    IGK_only    IGL_only 
#        851           3         439           6         149          94 

scBCR.metadata$IGH_expanded  %>% table() %>% as.data.frame()
scBCR.metadata$IGH_cloneSize %>% table() %>% as.data.frame()
scBCR.metadata$IGH_num       %>% table() %>% as.data.frame()
scBCR.metadata$IGK_expanded  %>% table() %>% as.data.frame()
scBCR.metadata$IGK_cloneSize %>% table() %>% as.data.frame()
scBCR.metadata$IGK_num       %>% table() %>% as.data.frame()
scBCR.metadata$IGL_expanded  %>% table() %>% as.data.frame()
scBCR.metadata$IGL_cloneSize %>% table() %>% as.data.frame()
scBCR.metadata$IGL_num       %>% table() %>% as.data.frame()
```

# --------------------------------------------

# merge scTCR + scBCR metadata

```{r}
scTCR.scBCR.metadata = 
  plyr::join_all(list(scTCR.metadata, scBCR.metadata), by="sample2_Barcode", type = "full") %>% 
  mutate(Tcell = if_else(is.na(Tcell), "noTCR", Tcell)) %>% 
  mutate(Bcell = if_else(is.na(Bcell), "noBCR", Bcell)) %>% 
  dplyr::select(sample2_Barcode, Tcell, Bcell, TRB_expanded, TRB_cloneSize, everything())
scTCR.scBCR.metadata
```

# saveRDS()

```{r}
# scTCR.scBCR.metadata %>% saveRDS(., file.path(dir.out.data, paste0(PROJ_BATCH, ".scTCR.scBCR.metadata.Rds")))
```

<br>

# EOF
