---
title: "2023_08_23_1011-IRIS_batch123_scGEX_load_crescent_DoubletFinder_annotate"
author: "Ming"
date: "`r format(Sys.time(), '%Y_%m_%d_%H_%M')`"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(eval = FALSE)      ## no run
# knitr::opts_chunk$set(include = FALSE)   ## yes run, no code, no output
knitr::opts_chunk$set(echo = TRUE)         ## yes run, no code, yes output ( for figures )
knitr::opts_chunk$set(results = 'hide')    ## yes run, yes code, no output
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
'%ni%' = Negate('%in%')
# options(scipen = 999)       # decimals
# options(scipen = 0)         # sci notation
# options(max.print=999999)

dir.ref                 = "/Users/minghan/bioinfoproj/references"
dir.clin                = "/Users/minghan/GDrive_minghanpughlab/PughLabPMH/_projects/IRIS/_clin"
dir.data                = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/data"
dir.out.data            = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_data"
dir.out.figure.analysis = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_analysis"
dir.out.figure.qc       = "/Users/minghan/bioinfoproj/PughLabPMH/_projects/IRIS/scmulti/out_figures_qc"

PROJ = "IRIS"
PROJ_BATCH = "IRIS_batch123"

colours_n40 = 
  c("dodgerblue2", "#E31A1C", "green4", "#6A3D9A", "#FF7F00", 
    "gold1", "gray70", "#C1645C", "skyblue2", "#FB9A99", "palegreen2", 
    "#CAB2D6", "#FDBF6F", "#D46A43", "khaki2", "maroon", 
    "orchid1", "deeppink1", "blue1", "steelblue4", "darkturquoise", 
    "green1", "yellow4", "yellow3", "darkorange4", "brown",
    "#5C9A5C", "#CC8BAD", "#B8782A", "#546C9D", "#459E6F",
    
    "chocolate4", "chartreuse4", "burlywood", "deeppink4", "lightslateblue",
    "firebrick", "mintcream", "mediumpurple3", "springgreen4", "slateblue4")

library(Seurat)
```

# read in data

## seurat.obj

```{r}
crescent.seurat.obj = readRDS(file.path(dir.out.data, "../out_crescent/IRIS_batch123_noIRS0070_newREF/crescent.SEURAT_DGE.rds"))
# crescent.seurat.obj
# 39388 features across 48374 samples within 3 assays
crescent.seurat.obj@meta.data$dataset %>% unique() %>% sort() %>% length() # cat(., sep="\n")
# [1] 11
```

### explore seurat obj

```{r}
crescent.seurat.obj@assays$RNA
# Assay data with 18433 features for 48374 cells
# First 10 features:
#  SAMD11, NOC2L, KLHL17, PLEKHN1, PERM1, HES4, ISG15, AGRN, C1orf159, TTLL10 

crescent.seurat.obj@assays$SCT
# SCTAssay data with 17955 features for 48374 cells, and 11 SCTModel(s) 
# First 10 features:
#  SAMD11, NOC2L, KLHL17, PLEKHN1, HES4, ISG15, AGRN, C1orf159, TTLL10, TNFRSF18 
crescent.seurat.obj@assays$SCT@data %>% min() # [1] 0
crescent.seurat.obj@assays$SCT@data %>% max() # [1] 9.401291
crescent.seurat.obj@assays$SCT@scale.data %>% min() # [1] -7.071058
crescent.seurat.obj@assays$SCT@scale.data %>% max() # [1] 19.33342

crescent.seurat.obj@assays$integrated
# SCTAssay data with 3000 features for 48374 cells, and 1 SCTModel(s) 
# Top 10 variable features:
#  S100A2, GNLY, CCL4, KRT5, GZMB, S100A8, COL1A1, KRT17, APOE, HLA-DRA 
crescent.seurat.obj@assays$integrated@SCTModel.list
crescent.seurat.obj@assays$integrated@var.features %>% length() # [1] 3000
crescent.seurat.obj@assays$integrated@data %>% min() # [1] -18.05736
crescent.seurat.obj@assays$integrated@data %>% max() # [1] 33.95997
crescent.seurat.obj@assays$integrated@scale.data %>% min() # [1] -17.95295
crescent.seurat.obj@assays$integrated@scale.data %>% max() # [1] 10

crescent.seurat.obj@active.assay # [1] "integrated"
crescent.seurat.obj@reductions
# $pca
# A dimensional reduction object with key PC_ 
#  Number of dimensions: 50 
#  Projected dimensional reduction calculated:  FALSE 
#  Jackstraw run: FALSE 
#  Computed using assay: integrated 
# 
# $umap
# A dimensional reduction object with key UMAP_ 
#  Number of dimensions: 2 
#  Projected dimensional reduction calculated:  FALSE 
#  Jackstraw run: FALSE 
#  Computed using assay: integrated 
# 
# $tsne
# A dimensional reduction object with key tSNE_ 
#  Number of dimensions: 2 
#  Projected dimensional reduction calculated:  FALSE 
#  Jackstraw run: FALSE 
#  Computed using assay: integrated 
```

### rename cells

```{r}
## get new cell vector
# crescent.seurat.obj@meta.data$dataset %>% unique() %>% sort()
crescent.seurat.obj.meta.data = crescent.seurat.obj@meta.data
crescent.seurat.obj.meta.data$dataset %>% unique() %>% cat(., sep = "\n")

crescent.seurat.obj.meta.data2 =
  crescent.seurat.obj.meta.data %>% 
  rownames_to_column(var="sample_barcode") %>% 
  mutate(dataset = gsub("_filtered_feature_bc_matrix", "", dataset),
         EachDatasetGlobalCellClusters = gsub("_filtered_feature_bc_matrix", "", EachDatasetGlobalCellClusters)) %>% 
  mutate(sample_barcode = gsub(".*_", "", sample_barcode),
         sample_barcode = paste0(dataset, "_", sample_barcode)) %>% 
  column_to_rownames(var="sample_barcode")
crescent.seurat.obj.meta.data2$dataset %>% unique() %>% cat(., sep = "\n")
crescent.seurat.obj.meta.data2

## rename cells
crescent.seurat.obj.cell_renamed = 
  Seurat::RenameCells(crescent.seurat.obj, new.names = rownames(crescent.seurat.obj.meta.data2))

## sanity check - cells renamed
# crescent.seurat.obj.cell_renamed@meta.data %>%
#   rownames_to_column(var="sample_barcode") %>%
#   dplyr::filter(str_detect(sample_barcode, "IRS_0025_A_025"))

rm(crescent.seurat.obj)
crescent.seurat.obj.cell_renamed

## rename other metadata columns
crescent.seurat.obj.cell_renamed.meta.data = crescent.seurat.obj.cell_renamed@meta.data
crescent.seurat.obj.cell_renamed.meta.data2 =
  crescent.seurat.obj.cell_renamed.meta.data %>% 
  mutate(dataset = gsub("_filtered_feature_bc_matrix", "", dataset),
         EachDatasetGlobalCellClusters = gsub("_filtered_feature_bc_matrix", "", EachDatasetGlobalCellClusters))
crescent.seurat.obj.cell_renamed@meta.data = crescent.seurat.obj.cell_renamed.meta.data2
crescent.seurat.obj.cell_renamed %>% dim() # [1] 17955 48374

## sanity check - other columns renamed
# crescent.seurat.obj.cell_renamed@meta.data$dataset %>% unique()
# crescent.seurat.obj.cell_renamed@meta.data %>% dplyr::filter(str_detect(dataset, "IRS_0025_A_025"))
```

# --------------------------------------------

# bio QC

```{r}
Seurat::DefaultAssay(crescent.seurat.obj.cell_renamed) = "SCT"
```

## QC metdata

```{r}
crescent.seurat.obj.cell_renamed = Seurat::PercentageFeatureSet(crescent.seurat.obj.cell_renamed, "^HB[^(P|S)]", col.name = "percent_hb")

feats = c("nFeature_RNA", "nCount_RNA", "mito.fraction", "ribo.fraction", "percent_hb")
Seurat::VlnPlot(crescent.seurat.obj.cell_renamed, group.by = "dataset", features = feats, pt.size = 0.01, ncol = 3) + Seurat::NoLegend()
## { mito% is a bit high for some samples, 
##   HB% is pretty low so no RBC contam }

Seurat::FeatureScatter(crescent.seurat.obj.cell_renamed, "nCount_RNA", "nFeature_RNA", group.by = "dataset", pt.size = 0.1)
## { corr 0.91, good }
```

## additional filtering

```{r}
crescent.seurat.obj.cell_renamed

filter_out_c = Seurat::WhichCells(crescent.seurat.obj.cell_renamed, expression = nFeature_RNA <= 200)
filter_out_c %>% length() # [1] 40
selected_c = Seurat::WhichCells(crescent.seurat.obj.cell_renamed, expression = nFeature_RNA > 200)

filter_out_f = rownames(crescent.seurat.obj.cell_renamed)[Matrix::rowSums(crescent.seurat.obj.cell_renamed) <= 3]
filter_out_f %>% length() # [1] 132
# filter_out_f %>% sort()
selected_f = rownames(crescent.seurat.obj.cell_renamed)[Matrix::rowSums(crescent.seurat.obj.cell_renamed) > 3]

crescent.seurat.obj.cell_renamed.filt = subset(crescent.seurat.obj.cell_renamed, features = selected_f, cells = selected_c)
crescent.seurat.obj.cell_renamed.filt %>% dim()
# [1] 17823 48334

rm(crescent.seurat.obj.cell_renamed)
```

## plot QC post-filter

```{r}
Seurat::VlnPlot(crescent.seurat.obj.cell_renamed.filt, group.by = "dataset", features = feats, pt.size = 0.01, ncol = 3) + Seurat::NoLegend()
```

## remove genes

```{r}
# Filter MALAT1
crescent.seurat.obj.cell_renamed.filt2 = crescent.seurat.obj.cell_renamed.filt[!grepl("MALAT1", rownames(crescent.seurat.obj.cell_renamed.filt)), ]
rm(crescent.seurat.obj.cell_renamed.filt)

# Filter Mitocondrial
crescent.seurat.obj.cell_renamed.filt2 = crescent.seurat.obj.cell_renamed.filt2[!grepl("^MT-", rownames(crescent.seurat.obj.cell_renamed.filt2)), ]

# Filter Ribossomal gene (optional if that is a problem on your data)
crescent.seurat.obj.cell_renamed.filt2 = crescent.seurat.obj.cell_renamed.filt2[!grepl('^RP[SL]', rownames(crescent.seurat.obj.cell_renamed.filt2)), ]

# Filter Hemoglobin gene (optional if that is a problem on your data)
crescent.seurat.obj.cell_renamed.filt2 = crescent.seurat.obj.cell_renamed.filt2[!grepl("^HB[^(P|S)]", rownames(crescent.seurat.obj.cell_renamed.filt2)), ]

crescent.seurat.obj.cell_renamed.filt2 %>% dim() # [1] 17704 48334
```

## find doublets

```{r}
library(DoubletFinder)
library(parallel)

set.seed(8)
options(mc.cores = detectCores() - 1)

crescent.seurat.obj.cell_renamed.filt2.split = Seurat::SplitObject(crescent.seurat.obj.cell_renamed.filt2, split.by = "dataset")
crescent.seurat.obj.cell_renamed.filt2.split[[1]]
crescent.seurat.obj.cell_renamed.filt2.split %>% names()
rm(crescent.seurat.obj.cell_renamed.filt2)

## get 10x multiplet rate table
sc_multiplet_rate.df = readxl::read_xlsx(file.path(dir.ref, "single_cell/single_cell_multiplet_rate_10X.xlsx"), sheet = "Sheet1")
model.multiplet = lm(multiplet_rate_pct ~ num_cell_recovered, data = sc_multiplet_rate.df)
model.multiplet %>% summary()

# loop through samples to find doublets
doubletfinder_res.list = list()
for (i in 1:length(crescent.seurat.obj.cell_renamed.filt2.split)) {
  
  # i = 2
  
  sample_name = names(crescent.seurat.obj.cell_renamed.filt2.split)[i]; sample_name
  print(paste0("Sample ", sample_name))
  
  # Pre-process seurat object with standard seurat workflow
  Seurat::DefaultAssay(crescent.seurat.obj.cell_renamed.filt2.split[[i]]) = "RNA"
  dataset_ = 
    crescent.seurat.obj.cell_renamed.filt2.split[[i]] %>% 
    # Seurat::SCTransform(., variable.features.n = 3000, vars.to.regress = "mito.fraction") %>% 
    Seurat::NormalizeData(., normalization.method = "LogNormalize") %>%
    Seurat::FindVariableFeatures(., selection.method = "vst", nfeatures = 3000) %>%
    Seurat::ScaleData(., vars.to.regress = "mito.fraction") %>%
    Seurat::RunPCA(., nfeatures.print = 10)
  # dataset_@assays$SCT
  
  # Find significant PCs
  stdv = dataset_[["pca"]]@stdev
  sum.stdv = sum(dataset_[["pca"]]@stdev)
  percent.stdv = (stdv / sum.stdv) * 100
  cumulative = cumsum(percent.stdv)
  co1 = which(cumulative > 90 & percent.stdv < 5)[1]
  co2 = sort(which((percent.stdv[1:length(percent.stdv) - 1] - 
                       percent.stdv[2:length(percent.stdv)]) > 0.1), 
              decreasing = T)[1] + 1
  min.pc = min(co1, co2)
  min.pc # [1] 20
  
  # finish pre-processing
  dataset_ = 
    dataset_ %>% 
    Seurat::RunUMAP(., dims = 1:min.pc) %>% 
    Seurat::FindNeighbors(., dims = 1:min.pc) %>%             
    Seurat::FindClusters(., resolution = 0.1)

  # pK identification (no ground-truth)
  # sweep.list = DoubletFinder::paramSweep_v3(dataset_, PCs = 1:min.pc, num.cores = detectCores() - 1, sct = T)
  ## pre-processing the data using NormalizeData instead of SCTransfrom. NormalizeData is MUCH less memory intensive
  sweep.list = DoubletFinder::paramSweep_v3(dataset_, PCs = 1:min.pc, num.cores = detectCores() - 1, sct = F)
  # sweep.list %>% names() %>% head(3) # [1] "pN_0.05_pK_0.005" "pN_0.05_pK_0.01"  "pN_0.05_pK_0.02" 
  sweep.stats = DoubletFinder::summarizeSweep(sweep.list)
  # sweep.stats %>% head(3) %>% print.data.frame()
  #     pN    pK    BCreal
  # 1 0.05 0.005 0.8738591
  # 2 0.05  0.01 0.7842234
  # 3 0.05  0.02 0.8121942
  bcmvn = find.pK(sweep.stats)
  # bcmvn %>% head(3) %>% print.data.frame()
  #   ParamID    pK    MeanBC        VarBC  BCmetric
  # 1       1 0.005 0.8298595 0.0013639140  608.4398
  # 2       2  0.01 0.8392603 0.0008658573  969.2824
  # 3       3  0.02 0.8504077 0.0005774987 1472.5708
  
  # ggplot(bcmvn, aes(pK, BCmetric)) + geom_point()
  
  # Optimal pK is the max of the bomodality coefficent (BCmvn) distribution
  bcmvn.max = bcmvn[which.max(bcmvn$BCmetric),]
  optimal.pk = bcmvn.max$pK
  optimal.pk = as.numeric(levels(optimal.pk))[optimal.pk]
  # optimal.pk # [1] 0.23
  
  ## Homotypic doublet proportion estimate
  annotations = dataset_@meta.data$seurat_clusters
  homotypic.prop = modelHomotypic(annotations)
  
  multiplet_rate = predict(model.multiplet, newdata = data.frame(num_cell_recovered = nrow(dataset_@meta.data))) %>% as.numeric()/100
  # multiplet_rate
  
  nExp.poi = round(multiplet_rate * nrow(dataset_@meta.data))
  # nExp.poi
  nExp.poi.adj = round(nExp.poi * (1 - homotypic.prop))
  # nExp.poi.adj # [1] 702
  
  # run DoubletFinder
  dataset_ = doubletFinder_v3(seu = dataset_, 
                               PCs = 1:min.pc,        ## statistically significant PCs, estimate from PC elbow
                               pN = 0.25,             ## proportion of real data to generate artificial doublets
                               pK = optimal.pk,       ## The PC neighborhood size used to compute pANN. Optimal pK values can be determined using mean-variance-normalized bimodality coefficient.
                               nExp = nExp.poi.adj)   ## total #of doublets predicted; value can best be estimated from cell loading densities
  # dataset_@meta.data
  metadata = 
    dataset_@meta.data %>% 
    dplyr::rename_with(.cols = starts_with("DF."), .fn = ~ gsub("DF.*", "doublet_finder", .x)) %>% 
    dplyr::rename_with(.cols = starts_with("pANN"), .fn = ~ gsub("pANN.*", "DoubletFinder_score", .x)) ## higher the score, more probable it's Doublet
  dataset_@meta.data = metadata
  # dataset_@meta.data$doublet_finder %>% table()
  # Doublet Singlet 
  #   702    3230 
  
  Seurat::DimPlot(dataset_, reduction = "umap", group.by = "doublet_finder") + labs(title = sample_name)
  
  # subset and save
  # dataset_singlets = subset(dataset_, doublet_finder == "Singlet")
  # dataset_singlets
  
  doubletfinder_res.list[[sample_name]] =
    dataset_@meta.data %>% 
    dplyr::select(doublet_finder, DoubletFinder_score)

}
doubletfinder_res.list[[3]]

doubletfinder_res.df = doubletfinder_res.list %>% bind_rows()
doubletfinder_res.df %>% saveRDS(., file.path(dir.out.data, paste0(PROJ_BATCH, ".doubletfinder_res.df.Rds")))

doubletfinder_res.df %>% 
  mutate(dataset = rownames(.) %>% gsub(".{17}$", "", .)) %>% 
  dplyr::select(dataset, doublet_finder) %>% table()
#                 doublet_finder
# dataset          Doublet Singlet
#   IRS_0004_A_003      66    3152
#   IRS_0014_P_014      15    1519
#   IRS_002_A_001       92    3840
#   IRS_0023_P_023     783   10346
#   IRS_0025_A_025       1     382
#   IRS_0032_A_032       4     907
#   IRS_0038_P_038     426    7925
#   IRS_0055_P_055     362    7461
#   IRS_0056_P_056       6     956
#   IRS_008_P_008        1     235
#   IRS_015_A_015      614    9241
```

## readRDS()

```{r}
doubletfinder_res.df = readRDS(file.path(dir.out.data, paste0(PROJ_BATCH, ".doubletfinder_res.df.Rds")))
doubletfinder_res.df %>% dplyr::filter(doublet_finder == "Singlet")
```

## remove doublets

```{r}
crescent.seurat.obj.cell_renamed.filt2.singlets = 
  subset(x = crescent.seurat.obj.cell_renamed.filt2, cells = doubletfinder_res.df %>% dplyr::filter(doublet_finder == "Singlet") %>% rownames())
crescent.seurat.obj.cell_renamed.filt2.singlets
```

## saveRDS()

```{r}
# crescent.seurat.obj.cell_renamed.filt2.singlets %>% saveRDS(., file.path(dir.out.data, paste0(PROJ_BATCH, ".crescent.seurat.obj.cell_renamed.filt2.singlets.Rds")))
crescent.seurat.obj.cell_renamed.filt2.singlets = readRDS(file.path(dir.out.data, paste0(PROJ_BATCH, ".crescent.seurat.obj.cell_renamed.filt2.singlets.Rds")))
```

# --------------------------------------------

# singleR 

## ENCODE reference

```{r}
ref.ENCODE = celldex::BlueprintEncodeData()
ENCODE_HN_cells.vec = c("Adipocytes", "B-cells", "CD4+ T-cells", "CD8+ T-cells", "DC", "Endothelial cells", "Eosinophils", "Epithelial cells", "Fibroblasts", "Macrophages", "Monocytes", "Neutrophils", "NK cells", "Skeletal muscle", "Smooth muscle")
ref.ENCODE.HN = ref.ENCODE[,ref.ENCODE$label.main %in% ENCODE_HN_cells.vec]
ref.ENCODE.HN %>% saveRDS(., file.path("/Users/minghan/bioinfoproj/references/single_cell/celldex/encode.hn.se.Rds"))
```

### run on H4H

- /cluster/projects/cgptip/projects/IRIS/projects/scmulti/analysis/analysis001_singleR/scripts

### load H4H results

```{r}
singleR.encode.res = readRDS(file.path(dir.out.data, "IRIS_batch123.singleR.encode_hn_mainLabel.Rds"))
singleR.encode.res.df = data.frame(cell_id = singleR.encode.res@rownames, 
                                   first.labels = singleR.encode.res$first.labels, 
                                   labels = singleR.encode.res$labels, 
                                   pruned.labels = singleR.encode.res$pruned.labels)
singleR.encode.res.df
```

```{r}
singleR.encode.res.df %>% mutate(sample_id = gsub('.{17}$', '', cell_id)) %>% dplyr::pull(sample_id) %>% table() %>% as.data.frame() %>% arrange(".") %>% print.data.frame()
#                 .  Freq
# 1  IRS_0004_A_003  3152
# 2  IRS_0014_P_014  1519
# 3   IRS_002_A_001  3840
# 4  IRS_0023_P_023 10346
# 5  IRS_0025_A_025   382
# 6  IRS_0032_A_032   907
# 7  IRS_0038_P_038  7925
# 8  IRS_0055_P_055  7461
# 9  IRS_0056_P_056   956
# 10  IRS_008_P_008   235
# 11  IRS_015_A_015  9241

singleR.encode.res.df$labels %>% table() %>% sum() # [1] 45964
singleR.encode.res.df$labels %>% table() %>% as.data.frame() %>% dplyr::rename("cell_type"=".") %>% arrange(desc(Freq)) %>% print.data.frame()
#            cell_type  Freq
# 1       CD8+ T-cells 13441
# 2        Fibroblasts 10167
# 3          Monocytes  6274
# 4        Macrophages  3439
# 5       CD4+ T-cells  3123
# 6  Endothelial cells  2474
# 7   Epithelial cells  1846
# 8                 DC  1676
# 9           NK cells  1534
# 10        Adipocytes   882
# 11           B-cells   748
# 12       Neutrophils   162
# 13     Smooth muscle   124
# 14       Eosinophils    73
# 15   Skeletal muscle     1
```

### plotScoreHeatmap()

```{r}
png(file.path(dir.out.figure.qc, "IRIS_batch123.singleR.encode_hn_mainLabel.plotScoreHeatmap.png"), 
    width = 8, height = 4, units = "in", bg = "white", res = 300)
SingleR::plotScoreHeatmap(singleR.encode.res)
dev.off()
```

### merge into seurat obj

```{r}
crescent.seurat.obj.singleR.encode = crescent.seurat.obj.cell_renamed.filt2.singlets
crescent.seurat.obj.singleR.encode.meta.data = crescent.seurat.obj.singleR.encode@meta.data
crescent.seurat.obj.singleR.encode@meta.data = 
  crescent.seurat.obj.singleR.encode.meta.data %>% 
  rownames_to_column(var="cell_barcode") %>% 
  left_join(., singleR.encode.res.df %>% dplyr::select(cell_id, labels), by=c("cell_barcode"="cell_id")) %>% 
  dplyr::rename("singleR.labels"="labels") %>% 
  column_to_rownames(var="cell_barcode")
```

### DimPlot()

```{r}
Seurat::DimPlot(crescent.seurat.obj.singleR.encode, 
                group.by = "singleR.labels", split.by = "dataset", ncol = 4, reduction = "umap") + 
  scale_color_manual(values = colours_n40[3:20])
ggsave(file.path(dir.out.figure.analysis, "IRIS.annotation_singleR.encode_ref.batch123_n11.png"), device = "png", width = 12, height = 6, units = "in", dpi = 300, bg = "white")
```

### add scVDJ metadata

```{r}
scTCR.scBCR.metadata = readRDS(file.path(dir.out.data, paste0(PROJ_BATCH, ".scTCR.scBCR.metadata.Rds")))
setdiff(scTCR.scBCR.metadata$sample2_Barcode, crescent.seurat.obj.singleR.encode@meta.data %>% rownames) %>% length() # [1] 2234

crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata = 
  full_join(
    crescent.seurat.obj.singleR.encode@meta.data %>% rownames_to_column(var="sample_barcode"),
    scTCR.scBCR.metadata,
    by = c("sample_barcode"="sample2_Barcode")
  )
crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata
```

#### inspect metadata

```{r}
crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata %>% 
  mutate(Tcell_Bcell = paste0(Tcell, "_", Bcell)) %>% 
  dplyr::select(singleR.labels, Tcell_Bcell) %>% 
  table(useNA = "ifany")
#                    Tcell_Bcell
# singleR.labels      NA_NA noTCR_yesBCR yesTCR_noBCR yesTCR_yesBCR
#   Adipocytes          719           92           69             2
#   B-cells             218          493            4            33
#   CD4+ T-cells        891            2         2223             7
#   CD8+ T-cells       2546           25        10842            28
#   DC                 1511            9          154             2
#   Endothelial cells  2397            3           72             2
#   Eosinophils          73            0            0             0
#   Epithelial cells   1785            5           56             0
#   Fibroblasts        9568          138          445            16
#   Macrophages        2939          136          340            24
#   Monocytes          5582          108          563            21
#   Neutrophils         161            0            1             0
#   NK cells           1355           15          164             0
#   Skeletal muscle       1            0            0             0
#   Smooth muscle       124            0            0             0
#   <NA>                  0          366         1853            15
```

### revise singleR.label

```{r}
crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata.revised =
  crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata %>% 
  mutate(Tcell_Bcell = paste0(Tcell, "_", Bcell)) %>% 
  mutate(singleR.labels.revised = 
           case_when(
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Adipocytes"        ~ "Adipocytes_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "B-cells"           ~ "B-cells_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "CD4+ T-cells"      ~ "CD4+ T-cells", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "CD8+ T-cells"      ~ "CD8+ T-cells", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "DC"                ~ "DC_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Endothelial cells" ~ "Endothelial cells_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Eosinophils"       ~ "Eosinophils_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Epithelial cells"  ~ "Epithelial cells_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Fibroblasts"       ~ "Fibroblasts_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Macrophages"       ~ "Macrophages_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Monocytes"         ~ "Monocytes_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Neutrophils"       ~ "Neutrophils_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "NK cells"          ~ "NK cells_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Skeletal muscle"   ~ "Skeletal muscle_TCR", 
             Tcell_Bcell == "yesTCR_noBCR" & singleR.labels == "Smooth muscle"     ~ "Smooth muscle_TCR", 
             
             ## noTCR_yesBCR
             Tcell_Bcell == "noTCR_yesBCR" &  str_detect(singleR.labels, "B-cells") ~ singleR.labels,
             Tcell_Bcell == "noTCR_yesBCR" & !str_detect(singleR.labels, "B-cells") ~ "B-cells",
             
             ## yesTCR_yesBCR
             Tcell_Bcell == "yesTCR_yesBCR" & !is.na(singleR.labels) ~ "TCR_BCR_DualExpressor",
             
             ## noTCR_noBCR
             Tcell_Bcell == "NA_NA" & singleR.labels == "CD4+ T-cells" ~ "CD4+ T-cells_noTCR",
             Tcell_Bcell == "NA_NA" & singleR.labels == "CD8+ T-cells" ~ "CD8+ T-cells_noTCR",
             Tcell_Bcell == "NA_NA" & str_detect(singleR.labels, "B-cells") ~ "B-cells_noBCR",
             Tcell_Bcell == "NA_NA" & !str_detect(singleR.labels, "T-cells") & !str_detect(singleR.labels, "B-cells") ~ singleR.labels,
             
             TRUE ~ "ERROR")) %>% 
  dplyr::select(sample_barcode, orig.ident, dataset, singleR.labels, singleR.labels.revised, Tcell_Bcell, everything())
crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata.revised$singleR.labels.revised %>% unique() %>% sort() %>% cat(., sep = "\n")
crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata.revised %>% 
  dplyr::filter(singleR.labels.revised == "ERROR") %>% nrow() # [1] 2234

crescent.seurat.obj.singleR.encode@meta.data = 
  crescent.seurat.obj.singleR.encode.meta.data.scVDJmetadata.revised %>% 
  dplyr::filter(!is.na(dataset)) %>% 
  column_to_rownames(var = "sample_barcode")
crescent.seurat.obj.singleR.encode@meta.data
```

### saveRDS()

```{r}
crescent.seurat.obj.singleR.encode %>% saveRDS(., file.path(dir.out.data, paste0(PROJ_BATCH, ".crescent.seurat.obj.singleR.encode.Rds")))
```

<br>

# EOF
